---
title: "Practical Machine Learning Course Project"
author: "Faisal Hoda"
date: "June 1, 2017"
output: 
  html_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction


## Load Data
```{r "load data and libraries", message=FALSE, warning=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(readr)


pml_training <- read_csv("./pml-training.csv")
pml_testing <- read_csv("./pml-testing.csv")

```


## Data Exploration
```{r "data exploration", warning=FALSE}
dim(pml_training)
dim(pml_testing)

str(pml_training)
```

There are some variables which need to be changed to the appropriate data type.
Also there are a lot of variables that have NAs. This will be addressed when we clean the data.


## Data Cleaning and Processing
```{r "data cleaning", warning=FALSE}

# Remove columns with NAs
training = pml_training[colSums(is.na(pml_training)) == 0]
dim(training)
testing = pml_testing[colSums(is.na(pml_training)) == 0]
dim(testing)


# Create a training data set and a testing set.
set.seed(20170602)
inTrain <- createDataPartition(training$classe, p=0.6, list=FALSE)
model_training <- training[inTrain, ]
model_testing <- training[-inTrain, ]

dim(model_training)
dim(model_testing)
```


## Modeling 

### Random Forest

Generate a random forest model with cross validation to avoid overfitting. 

```{r "random forest", cache=TRUE}
modFitrf <- train(classe ~. - user_name - X1, method="rf", data=model_training, trControl=trainControl(method='cv'), number=3, allowParallel=TRUE, importance=TRUE )
modFitrf

```

Test Model and Confusion Matrix
```{r "RF test and confusion matrix"}
predTest <- predict(modFitrf, model_testing)
rf_cf <- confusionMatrix(predTest ,model_testing$classe)
rf_cf
```


### Decision Tree
```{r "decision tree", cache=TRUE}
modFitdt <- train(classe ~. - user_name - X1, method="rpart", data=model_training)

modFitdt

```

Test model and Confusion Matrix
```{r "DT test and consution matrix"}
predTest_dt <- predict(modFitdt, model_testing)
dt_cf <- confusionMatrix(predTest_dt ,model_testing$classe)
dt_cf

```


## Prediction on Test Data

The random forest model is the most accurate at $0.9983$, so that is the model being used for prediction.

```{r prediction}
final_prediction <- predict(modFitrf, testing)
final_prediction

```





